An efficient composition of bidirectional programs by tupling and lazy updates

### 2019-11-29 paper/5.experiments.tex
update references
add labels to sections

\section{Experiments}

@@

\subsection{Test cases}

We have conducted 7 test cases shown in Table \ref{tab:test-cases}. Although the number of programs is not so much, they are not trivial and can be completely used in practice. In $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and view $v_0$. This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is present in the table. These $bx$ are left associative composition since we we are only interested in this kind of program in the paper. Please note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.

@@

We introduced the composition $phead \circ phead$ earlier. The compositions of many \textit{phead} do things similarly, the head of a head element inside a super nested list which expresses the source should be changed by the view. Because of the type of the source, this program is seen as a \textit{ldata} case. Next, we briefly explain the behavior of the remaining compositions in table \ref{tab:test-cases}.

>>>>

\section{Experiments} \label{sect:experiments}

@@

\subsection{Test Cases}

We have conducted 7 test cases shown in Table~\ref{tab:test-cases}. Although the number of programs is not so much, they are not trivial and can be completely used in practice. In $s_r$ and $v_r$ are respectively updated source and view which are produced by applying $put$ and $get$ on original source $s_0$ and view $v_0$. This means: $s_r = \putbx{bx}{s_0}{v_0}$ and $v_r = \getbx{bx}{s_0}$, where $bx$ is present in the table. These $bx$ are left associative composition since we we are only interested in this kind of program in the paper. Please note that the results $s_r$ and $v_r$ do not depend on the associative style of the composition.

@@

We introduced the composition $phead \circ phead$ earlier. The compositions of many \textit{phead} do things similarly, the head of a head element inside a super nested list which expresses the source should be changed by the view. Because of the type of the source, this program is seen as a \textit{ldata} case. Next, we briefly explain the behavior of the remaining compositions in Table~\ref{tab:test-cases}.

### 2019-11-29 paper/6.related-work.tex
add labels to sections

\section{Related Work}

@@

There are many BX languages \cite{Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}
%(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)
based on several application scenarios and BX lenses works \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487}. The semantics and correctness of bidirectional programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood. Anjorin et. al. introduces the first benchmark for BX languages and compared them \cite{Anjorin2019}. However, the improvement for practical implementation of BX languages is still missing. As a first step, we study different evaluation methods to optimized their evaluation. We can say this paper will be first attempt to improve efficiency of BX composition evaluation. In this paper we focus on a BX language BiGUL \cite{Ko:2016:BFV:2847538.2847544, Ko:2017:ABB:3177123.3158129}, we compare BiGUL's method (in Section 2) and our methods (in Section 3 and 4). From experimental results of left associative BX composition programs, we know that our memoization methods are faster than the original BiGUL's evaluation method.
In this paper we focus on BiGUL, however our methods can be applicable to other BX languages which have similar problems.

@@

Memoization \cite{Bellman:2003:DP:862270,MICHIE1968} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
Another way to reduce extra evaluation is fusion \cite{Wadler:1988:DTP:80099.80104}. However, fusion of BX programs are not considered well \textcolor{red}{(really?)}. Therefore we did not use this way.
  Instead of that, we introduce another way that uses tupling \cite{Fokkinga90}, lazy update and lazy computation \cite{Henderson:1976:LE:800168.811543, Hudak:2007:HHL:1238844.1238856}. Tupling is often used for reducing common evaluations, but we use it for reducing the size of complements.
  Lazy evaluation is also a well-known and widly-used optimization technique to reduce the redundant evaluation. Because we use OCaml for implementation, we need to introduce this explicitly. If we use Haskell, a lazy evaluated language, we do not need to introduce them.

>>>>

\section{Related Work} \label{sect:related}

@@

There are many BX languages \cite{Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}
%(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)
based on several application scenarios and BX lenses works \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487}. The semantics and correctness of BX programs has been investigated intensively during the past years, but their efficiency and optimization are not yet fully understood. Anjorin et. al. introduces the first benchmark for BX languages and compared them \cite{Anjorin2019}. However, the improvement for practical implementation of BX languages is still missing. As a first step, we study different evaluation methods to optimize their evaluation. We can say this paper will be first attempt to improve efficiency of BX composition evaluation. In this paper we focus on a BX language BiGUL \cite{Ko:2016:BFV:2847538.2847544, Ko:2017:ABB:3177123.3158129}, we compare BiGUL's method (in Section~\ref{sect:minbigul}) and our methods (in Sections~\ref{sect:minbigulm} and \ref{sect:xpg}). From experimental results of left associative BX composition programs, we know that our memoization methods are faster than the original BiGUL's evaluation method.
In this paper we focus on a BX language, BiGUL, however our methods can be applicable to other BX languages which have similar problems.

@@

Memoization \cite{Bellman:2003:DP:862270,MICHIE1968} is well-known technique to reduce extra evaluation. In our case, simple application of memoization $put_m$ works well for many BX programs.
Another way to reduce extra evaluation is program fusion \cite{Wadler:1988:DTP:80099.80104}. However, we can not apply fusion to dynamically produced programs by recursion. Therefore we did not use fusion in this work.
  Instead of that, we introduced another way that uses tupling \cite{Fokkinga90}, lazy update and lazy computation \cite{Henderson:1976:LE:800168.811543, Hudak:2007:HHL:1238844.1238856}. Tupling is often used for reducing common evaluations, but we use it for reducing the size of complements.
  Lazy evaluation is also a well-known and widly-used optimization technique to reduce the redundant evaluation. Because we use OCaml for implementation, we need to introduce this explicitly. If we use Haskell, a lazy evaluated language, we do not need to introduce it.
